# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd("C:/Users/okamal.MGC/Documents/GitHub/RepData_PeerAssessment1")
raw_data <- read.csv("./activity.csv")
raw_data$date <- as.Date(raw_data$date,format="%Y-%m-%d")
```



## What is mean total number of steps taken per day?
```{r}
table_means <- aggregate(raw_data$steps,list(raw_data$date),FUN=mean)
table_median <- aggregate(raw_data$steps,list(raw_data$date),FUN=median)
table_sum_steps_per_day <-  aggregate(raw_data$steps,list(raw_data$date),FUN=sum)
hist(table_sum_steps_per_day$x,main="Histogram of sum of steps per day",xlab="Steps per day")
final_table <- data.frame(date=table_means$Group.1,mean=table_means$x,median=table_median$x,sum_of_steps=table_sum_steps_per_day$x)
final_table
```



## What is the average daily activity pattern?
```{r}
interval_means <- aggregate(raw_data$steps,list(raw_data$interval),FUN=mean,na.rm=T)
plot(y=interval_means$x,type="l",x=interval_means$Group.1,xlab="Interval",ylab="Average no of steps",main="Average Steps per interval across all Days")

interval_max <- interval_means[which.max(interval_means$x),1]

```

The 5-minute interval, that have the __highest average__ across all the days in the dataset, is interval __`r interval_max`__

## Imputing missing values
```{r}
no_missing_values <- sum(is.na(raw_data$steps))
# filling missing intervals with mean
imputed_data_list <- apply(raw_data,1,function(x) {
  if(is.na(x[1])) {
    x[1] <- interval_means[as.integer(interval_means$Group.1)==as.integer(x[3]),2]
  }
  return(as.vector(x))
  } )
imputed_data_frame <- as.data.frame(apply(imputed_data_list,1,rbind),stringsAsFactors=F)
colnames(imputed_data_frame) <- c("steps","date","interval")
imputed_data_frame$steps <- as.numeric(imputed_data_frame$steps)
imputed_data_frame$date <- as.Date(imputed_data_frame$date,format="%Y-%m-%d")
imputed_data_frame$interval <- as.integer(imputed_data_frame$interval)

imputed_sum_steps_per_day <-  aggregate(imputed_data_frame$steps,list(imputed_data_frame$date),FUN=sum)
hist(imputed_sum_steps_per_day$x,main="Histogram for sum of steps per data for imputed data",xlab="sum of steps per day")

imputed_means <- aggregate(imputed_data_frame$steps,list(imputed_data_frame$date),FUN=mean)
imputed_median <- aggregate(imputed_data_frame$steps,list(imputed_data_frame$date),FUN=median)


nrows_imputed_data <- dim(imputed_means)[1]
type_data <- rep(c("imputed","original"),each=nrows_imputed_data)
total_mean_data <- cbind(rbind(imputed_means,table_means),type=type_data)
total_median_data <- cbind(rbind(imputed_median,table_median),type=type_data)


  
```



Here is the histograms for the imputed data 

```{r}

require(ggplot2)
ggplot(imputed_sum_steps_per_day, aes(x=x)) + geom_histogram(binwidth=1000, alpha=.5, position="identity") + labs(title="Histogram for imputed sum of steps per day")
ggplot(total_mean_data, aes(x=x, fill=type,colour=type)) + geom_histogram(binwidth=3, alpha=.5, position="identity") + labs(title="Histogram for original and imputed means")
ggplot(total_median_data, aes(x=x, fill=type,colour=type)) + geom_histogram(binwidth=3, alpha=.5, position="identity") + labs(title="Histogram original and imputed medians")
     
```


Here is the imputed mean and median total number of steps taken per day.
### Imputed mean steps per day

```{r}
imputed_means
```

### Imputed median steps per day

```{r}
imputed_median
```

Yes there is a difference in some days between the original data and the imputed data.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
imputed_means_with_weekdata <- imputed_data_frame
imputed_means_with_weekdata$weekday <- weekdays(imputed_data_frame$date)
imputed_means_with_weekdata$day_type <- ifelse(imputed_means_with_weekdata$weekday %in% c("Saturday","Sunday"), "Weekend","Weekday") 
mean_steps_weekdays <- aggregate(imputed_means_with_weekdata$steps,list(imputed_means_with_weekdata$interval,imputed_means_with_weekdata$day_type),FUN=mean)
colnames(mean_steps_weekdays) <- c("interval","day_type","mean_steps")

require(lattice)
xyplot(mean_steps~interval|day_type,data=mean_steps_weekdays,layout=c(1,2),type="l",ylab="Mean number of steps")
```

